PUT YOUR TITLE HERE

### 2023-03-22 paper/manuscript.Rmd
tweaks

[replication: it's a hot issue and concerns around replication have done a lot, see crisis/revolution] Concerns around the practices of science. Replication is a hot topic. Some argue that it is a cornerstone of a cumulative science, and findings of low replication rates are a problem. Concerns about replication rates have become a large point of discussion in psychology and other fields, with large scale replications spurring these discussions and providing the empirical data analyzed by all sides.

[ there's ... not a lot of empirics]To quantify and understand this problem, we need empirical data about replication attempts. However, due to the arduous nature of collecting samples of replications, there have not been many large-scale replication efforts, so all of the argumentation around predictors of replications and field-wide replication rates is fit to a small number of data points.

Many replications of individual effects have been performed, but these are less useful as pattern analysis because the varied sampling and reporting, and the fact they aren't all in one place. Also pub bias.

We are aware of three large-scale replication efforts replicating experimental results in the existing psychology literature. RP:P sampled roughly 100 studies from top psychology journals in 2008 [@openscienceconsortium2015]. They found an overall replication rate around 40%, which provided evidence to support growing concerns about the non-reliability of the literature. The RP:P dataset was itself much discussed and reanalyzed [@etz2016;@gilbert2016;@patil2016] [idk, maybe worth framing that this was a big deal study when it came out]. TODO some discussion of how many papers have reanalysed this cutting the data different ways.

@@

In addition to determining estimated overall replicaiton rates for fields and journals, there's also merit in knowing what features of experiments (and replication attempts) are predictive of replication success. Blah blah stakeholders and resources. There's some signal here, as people are able to predict replication success at above chance CITATIONS [@hoogeveen2019;@dreber2015; @forsell2019;@camerer2018]

@@

Prior literature, such as @frank2012 and @hawkins have advocated using research methods classes to conduct replications, both for the education and scientific value.

@@

PSYCH 251 is Stanford Psychology's graduate-level experimental methods class taught by MCF.  Over the course of 10 weeks, students choose a published finding to replicate, re-implement the study, write analysis code, pre-register their study, collect data using an online platform, and write up a structured replication report. Students are free to choose studies related to their research interests, with the default recommendation being an article from a recent year of Psychological Science. While this choice results in a non-random sampling of studies, the sample is representative of studies that are of interest to and doable by first year graduate students.

The sample of replicated studies reflects the variability of the literature, including studies from different subfields, using different experimental methods and statistical outcomes. We leverage the naturally occuraing variability in this sample of replications to examine how different demographic, experimental design, and statistical properties predict replication success.

We used a subjective rating of replication success as our primary outcome measure. The instruction team had coded a holistic measure of replication success for each project when they were turned in at the end of the course. For reliability, we had a separate rater (VB) independently code the replication success from the replication reports; discrepancies were resolved by discussion between MCF and VB. We focussed on this measure because it was applicable across the range of statistical measures and reporting practices and accommodated studies where there were multiple important outcome measures.

@@

pred_int_rate <- d |> filter(!is.na(predInt)) |> summarize(m=mean(as.numeric(predInt))*100)
porig_rate <- d |> filter(!is.na(p_orig)) |> summarize(m=median(p_orig)*100)
```
Across the 176 studies, the overall subjective replication rate was `r round(pluck(sub_rep_rate,1))`%. `r round(pluck(pred_int_rate,1))`% of the studies had replication outcomes within the prediction interval of the original outcome. The median p_original value was `r round(pluck(porig_rate, 1))`%. Figure \ref{fig:smd} shows the relationship between original SMD, replication SMD, and subjective replication score. Roughly speaking, there's a cluster of studies that replicate with similar effect sizes to the original and another cluster that fail to replicate with replication effect sizes near zero. On average, there is a diminution of effect sizes from original to replication.

@@

## Dataset
The dataset of replication projects comes from class projects conducted in PSYCH 251 (earlier called PSYCH 24) a graduate-level experimental methods class taught by MCF from 2011 to 2022. Each student chose a study to replicate, implemented the study, wrote analysis code, pre-registered their replication, ran the study, and turned in a structured final report including methods, analytic plan, changes from the original study, confirmatory and exploratory analyses, and discussion of outcomes. Students were encouraged to do experimental replications, but some students chose to replicate correlational outcomes or do computational reproducibility projects instead. We cannot include the full student reports for confidentiality reasons, but we include an example as well as the template given to students in the repo. TODO example and template

[some of this should actually just go in methods methods] Class, taught by last author, is an experimental methods class aimed at first-year graduate students, but taken by ... Since around 2015, it has been required for incoming PhD students in the Psychology Department.Students are free to choose studies related to their interests, although recent articles from Psychological Science are the recommended path under uncertainty (leading to a high proportion of Psych science articles in this replication sample). While the sampling procedure is non-random, it is generally representative of studies that are of interest to and doable by first year grad students.

>>>>

[structure w/i intro?]

[replication: it's a hot issue and concerns around replication have done a lot, see crisis/revolution] Psychology is in the middle of some heated debate around the fields practices and many reform efforts. Many of these issues concern the truthworthiness or not of the published literature, with transparency, openness efforts. A lot of this centers around issues of reproducibility and replicability.  Concerns around the practices of science. Replication is a hot topic. Some argue that it is a cornerstone of a cumulative science, and findings of low replication rates are a problem. Concerns about replication rates have become a large point of discussion in psychology and other fields, with large scale replications spurring these discussions and providing the empirical data analyzed by all sides.

[ there's ... not a lot of empirics] One of the concerns is that published literature may not replicate as much as previously thought or as desired. To quantify and understand this problem, we need empirical data about replication attempts. However, due to the arduous nature of collecting samples of replications, there have not been many large-scale replication efforts, so all of the argumentation around predictors of replications and field-wide replication rates is fit to a small number of data points.

[do we discuss reasons for replications and why many, while valuable in specific phenomena, help less with this question?]
Many replications of individual effects have been performed, but these are less useful as pattern analysis because the varied sampling and reporting, and the fact they aren't all in one place. Also pub bias.

## prior literature
We are aware of three large-scale replication efforts replicating experimental results in the existing psychology literature. RP:P sampled roughly 100 studies from top psychology journals in 2008 [@openscienceconsortium2015]. They found an overall replication rate around 40%, which provided evidence to support growing concerns about the non-reliability of the literature. The RP:P dataset was itself much discussed and reanalyzed [@etz2016;@gilbert2016;@patil2016] [idk, maybe worth framing that this was a big deal study when it came out]. TODO some discussion of how many papers have reanalysed this cutting the data different ways.

@@

## predictors
In addition to determining estimated overall replicaiton rates for fields and journals, there's also merit in knowing what features of experiments (and replication attempts) are predictive of replication success. Blah blah stakeholders and resources. Scientists may want to build their work on studies that are likely to replicate, so that they can replicate and build on work without wasting resources. Similarly investing in interventions may want to start with things that are more likely to replicate. There's some signal here, as people are able to predict replication success at above chance CITATIONS [@hoogeveen2019;@dreber2015; @forsell2019;@camerer2018]

@@

## current work
Prior literature, such as @frank2012 and @hawkins have advocated using research methods classes to conduct replications, both for the education and scientific value.

@@

PSYCH 251 is Stanford Psychology's graduate-level experimental methods class taught by MCF.  During the 10 week class, students replicate a published finding. They individually re-implement the study, write analysis code, pre-register their study, collect data using an online platform, and write up a structured replication report. Students are free to choose studies related to their research interests, with the default recommendation being an article from a recent year of Psychological Science. While this choice results in a non-random sample from the literature, the sample is representative of studies that are of interest to and doable by first year graduate students.

The sample of replicated studies reflects the variability of the literature, including studies from different subfields, using different experimental methods and statistical outcomes. We leverage the naturally occurring variability in this sample of replications to examine how different demographic, experimental design, and statistical properties predict replication success.

We used a subjective rating of replication success as our primary outcome measure. The instruction team had coded a holistic measure of replication success for each project when they were turned in at the end of the course. For reliability, VB independently code the replication success from the replication reports; discrepancies were resolved by discussion between MCF and VB (ToDO foobar % of cases). This measure was applicable across the range of statistical measures and reporting practices and accommodated studies where there were multiple important outcome measures.

@@

pred_int_rate <- d |> filter(!is.na(predInt)) |> summarize(m=mean(as.numeric(predInt))*100)
porig_rate <- d |> filter(!is.na(p_orig)) |> summarize(m=median(p_orig))
```

Across the 176 studies, the overall subjective replication rate was `r round(pluck(sub_rep_rate,1))`%. `r round(pluck(pred_int_rate,1))`% of the studies had replication outcomes within the prediction interval of the original outcome. The median p_original value was `r round(pluck(porig_rate, 1),2)`. Figure \ref{fig:smd} shows the relationship between original SMD, replication SMD, and subjective replication score. Roughly speaking, there's a cluster of studies that replicate with similar effect sizes to the original and another cluster that fail to replicate with replication effect sizes near zero. On average, there is a diminution of effect sizes from original to replication.

@@

## Dataset
The dataset of replication projects comes from class projects conducted in PSYCH 251 (earlier called PSYCH 254) a graduate-level experimental methods class taught by MCF from 2011 to 2022. This class is commonly taken by first year graduate students in psychology and related disciplines, and it has been a requirement of the Psychology PhD since around 2015. Each student chose a study to replicate, implemented the study, wrote analysis code, pre-registered their replication, ran the study, and turned in a structured final report including methods, analytic plan, changes from the original study, confirmatory and exploratory analyses, and discussion of outcomes. Students were encouraged to do experimental replications, but some students chose to replicate correlational outcomes or do computational reproducibility projects instead. We cannot include the full student reports for confidentiality reasons, but we include an example as well as the template given to students in the repo. TODO example and template

Students were free to choose what study they replicated; the recommended path for students who did not have their own ideas was to pick an interesting study from a recent year of Psychological Science (this led to a high fraction of Psych Science articles in the replication sample FOOBAR %).

### 2023-03-22 paper/manuscript.tex
tweaks

{[}replication: it's a hot issue and concerns around replication have done a lot, see crisis/revolution{]} Concerns around the practices of science. Replication is a hot topic. Some argue that it is a cornerstone of a cumulative science, and findings of low replication rates are a problem. Concerns about replication rates have become a large point of discussion in psychology and other fields, with large scale replications spurring these discussions and providing the empirical data analyzed by all sides.

{[} there's \ldots{} not a lot of empirics{]}To quantify and understand this problem, we need empirical data about replication attempts. However, due to the arduous nature of collecting samples of replications, there have not been many large-scale replication efforts, so all of the argumentation around predictors of replications and field-wide replication rates is fit to a small number of data points.

Many replications of individual effects have been performed, but these are less useful as pattern analysis because the varied sampling and reporting, and the fact they aren't all in one place. Also pub bias.

@@

In addition to determining estimated overall replicaiton rates for fields and journals, there's also merit in knowing what features of experiments (and replication attempts) are predictive of replication success. Blah blah stakeholders and resources. There's some signal here, as people are able to predict replication success at above chance CITATIONS (\protect\hyperlink{ref-dreber2015}{Dreber et al. 2015}, \protect\hyperlink{ref-camerer2018}{Camerer et al. 2018}, \protect\hyperlink{ref-forsell2019}{Forsell et al. 2019}, \protect\hyperlink{ref-hoogeveen2019}{Hoogeveen et al. 2019})

@@

@@

PSYCH 251 is Stanford Psychology's graduate-level experimental methods class taught by MCF. Over the course of 10 weeks, students choose a published finding to replicate, re-implement the study, write analysis code, pre-register their study, collect data using an online platform, and write up a structured replication report. Students are free to choose studies related to their research interests, with the default recommendation being an article from a recent year of Psychological Science. While this choice results in a non-random sampling of studies, the sample is representative of studies that are of interest to and doable by first year graduate students.

The sample of replicated studies reflects the variability of the literature, including studies from different subfields, using different experimental methods and statistical outcomes. We leverage the naturally occuraing variability in this sample of replications to examine how different demographic, experimental design, and statistical properties predict replication success.

We used a subjective rating of replication success as our primary outcome measure. The instruction team had coded a holistic measure of replication success for each project when they were turned in at the end of the course. For reliability, we had a separate rater (VB) independently code the replication success from the replication reports; discrepancies were resolved by discussion between MCF and VB. We focussed on this measure because it was applicable across the range of statistical measures and reporting practices and accommodated studies where there were multiple important outcome measures.

@@

Across the 176 studies, the overall subjective replication rate was 49\%. 46\% of the studies had replication outcomes within the prediction interval of the original outcome. The median p\_original value was 3\%. Figure \ref{fig:smd} shows the relationship between original SMD, replication SMD, and subjective replication score. Roughly speaking, there's a cluster of studies that replicate with similar effect sizes to the original and another cluster that fail to replicate with replication effect sizes near zero. On average, there is a diminution of effect sizes from original to replication.

@@

The dataset of replication projects comes from class projects conducted in PSYCH 251 (earlier called PSYCH 24) a graduate-level experimental methods class taught by MCF from 2011 to 2022. Each student chose a study to replicate, implemented the study, wrote analysis code, pre-registered their replication, ran the study, and turned in a structured final report including methods, analytic plan, changes from the original study, confirmatory and exploratory analyses, and discussion of outcomes. Students were encouraged to do experimental replications, but some students chose to replicate correlational outcomes or do computational reproducibility projects instead. We cannot include the full student reports for confidentiality reasons, but we include an example as well as the template given to students in the repo. TODO example and template

{[}some of this should actually just go in methods methods{]} Class, taught by last author, is an experimental methods class aimed at first-year graduate students, but taken by \ldots{} Since around 2015, it has been required for incoming PhD students in the Psychology Department.Students are free to choose studies related to their interests, although recent articles from Psychological Science are the recommended path under uncertainty (leading to a high proportion of Psych science articles in this replication sample). While the sampling procedure is non-random, it is generally representative of studies that are of interest to and doable by first year grad students.

>>>>

{[}structure w/i intro?{]}

{[}replication: it's a hot issue and concerns around replication have done a lot, see crisis/revolution{]} Psychology is in the middle of some heated debate around the fields practices and many reform efforts. Many of these issues concern the truthworthiness or not of the published literature, with transparency, openness efforts. A lot of this centers around issues of reproducibility and replicability. Concerns around the practices of science. Replication is a hot topic. Some argue that it is a cornerstone of a cumulative science, and findings of low replication rates are a problem. Concerns about replication rates have become a large point of discussion in psychology and other fields, with large scale replications spurring these discussions and providing the empirical data analyzed by all sides.

{[} there's \ldots{} not a lot of empirics{]} One of the concerns is that published literature may not replicate as much as previously thought or as desired. To quantify and understand this problem, we need empirical data about replication attempts. However, due to the arduous nature of collecting samples of replications, there have not been many large-scale replication efforts, so all of the argumentation around predictors of replications and field-wide replication rates is fit to a small number of data points.

{[}do we discuss reasons for replications and why many, while valuable in specific phenomena, help less with this question?{]}
Many replications of individual effects have been performed, but these are less useful as pattern analysis because the varied sampling and reporting, and the fact they aren't all in one place. Also pub bias.

\hypertarget{prior-literature}{%
\subsection{prior literature}\label{prior-literature}}

@@

\hypertarget{predictors}{%
\subsection{predictors}\label{predictors}}

In addition to determining estimated overall replicaiton rates for fields and journals, there's also merit in knowing what features of experiments (and replication attempts) are predictive of replication success. Blah blah stakeholders and resources. Scientists may want to build their work on studies that are likely to replicate, so that they can replicate and build on work without wasting resources. Similarly investing in interventions may want to start with things that are more likely to replicate. There's some signal here, as people are able to predict replication success at above chance CITATIONS (\protect\hyperlink{ref-dreber2015}{Dreber et al. 2015}, \protect\hyperlink{ref-camerer2018}{Camerer et al. 2018}, \protect\hyperlink{ref-forsell2019}{Forsell et al. 2019}, \protect\hyperlink{ref-hoogeveen2019}{Hoogeveen et al. 2019})

@@

\hypertarget{current-work}{%
\subsection{current work}\label{current-work}}

@@

PSYCH 251 is Stanford Psychology's graduate-level experimental methods class taught by MCF. During the 10 week class, students replicate a published finding. They individually re-implement the study, write analysis code, pre-register their study, collect data using an online platform, and write up a structured replication report. Students are free to choose studies related to their research interests, with the default recommendation being an article from a recent year of Psychological Science. While this choice results in a non-random sample from the literature, the sample is representative of studies that are of interest to and doable by first year graduate students.

The sample of replicated studies reflects the variability of the literature, including studies from different subfields, using different experimental methods and statistical outcomes. We leverage the naturally occurring variability in this sample of replications to examine how different demographic, experimental design, and statistical properties predict replication success.

We used a subjective rating of replication success as our primary outcome measure. The instruction team had coded a holistic measure of replication success for each project when they were turned in at the end of the course. For reliability, VB independently code the replication success from the replication reports; discrepancies were resolved by discussion between MCF and VB (ToDO foobar \% of cases). This measure was applicable across the range of statistical measures and reporting practices and accommodated studies where there were multiple important outcome measures.

@@

Across the 176 studies, the overall subjective replication rate was 49\%. 46\% of the studies had replication outcomes within the prediction interval of the original outcome. The median p\_original value was 0.03. Figure \ref{fig:smd} shows the relationship between original SMD, replication SMD, and subjective replication score. Roughly speaking, there's a cluster of studies that replicate with similar effect sizes to the original and another cluster that fail to replicate with replication effect sizes near zero. On average, there is a diminution of effect sizes from original to replication.

@@

The dataset of replication projects comes from class projects conducted in PSYCH 251 (earlier called PSYCH 254) a graduate-level experimental methods class taught by MCF from 2011 to 2022. This class is commonly taken by first year graduate students in psychology and related disciplines, and it has been a requirement of the Psychology PhD since around 2015. Each student chose a study to replicate, implemented the study, wrote analysis code, pre-registered their replication, ran the study, and turned in a structured final report including methods, analytic plan, changes from the original study, confirmatory and exploratory analyses, and discussion of outcomes. Students were encouraged to do experimental replications, but some students chose to replicate correlational outcomes or do computational reproducibility projects instead. We cannot include the full student reports for confidentiality reasons, but we include an example as well as the template given to students in the repo. TODO example and template

Students were free to choose what study they replicated; the recommended path for students who did not have their own ideas was to pick an interesting study from a recent year of Psychological Science (this led to a high fraction of Psych Science articles in the replication sample FOOBAR \%).
